Effects of sexual selection on non-sexual fitness components: a meta-analysis

### 2018-03-01 meta_analysis_sexual_selection_pop_fitness.Rmd
Luke and Justin meeting

```{r warning=FALSE, message=FALSE, results='hide'}
library(knitr)
library(pander)
library(compute.es)
library(metafor)
library(dplyr)
library(lme4)
library(forestplot)
library(ggplot2)
library(MuMIn)
library(glmulti)
library(cowplot)
library(ggrepel)
library(reshape2)
source("Vdodge_function.R") # nice function for ggplot
```

@@

**Study ID**: An ID given to the published paper the effect size is sourced from
**Group ID**: An ID given to the research group that may have published several papers on the same species usuing the same or very similar experimental setup
**Species and Species ID**: Same thing
**SS Strength, Ratios and SS Density's (Column 7-9)**: Various ratios of the number of males to females and the total number of individuals kept together in an experiment
**Ratio Category**: A three level category for the ratio of males to females (high, medium, low).
**Density Category**: A three level category for the density of males to females (high, medium, low).
**SSS.Categorical**: A three level category for the **density & the ratio** of males to females (high medium, low). 
**Post cop and Pre cop**: Whether a study allowed Pre/Post-copulatory sexual selection (1) or not (0)
**Blind**: Whether the study was blind or not
**Generations** The number of generations run before fitness outcomes were measured
**Enforced Monogamy**: Whether the study had the low sexual selection treatment as enforced monogamy (YES) or not (NO). Not all studies compared enforced monogamy and SS+ treatments. Some used FB vs MB, where FB is the SS (low intensity). 
**Sex**: Whether the fitness outcome was measured for females (F), males (M) or both (B). Studies that reported 'both' would pool results from males and females.
**Ambiguous**: Is the fitness outcome ambiguous (YES) or not ambigous (NO). Ambiguous outcomes may be those that may not necessarily be directional, that is to say they may be a life history trait. 
**Outcome Class**: Grouped as Direct, Indirect or Ambiguous.
**Environment**: In the methods of the papers included in this study it was usually stated whether additional modifications to the experimental lines were made. Briefly, this was usually a modification that made conditions more stressful such as using a novel food source or elevated mutation load, the effect sizes from these experimental lines are labelled as 'Stressed'. If it was clearly stated that there was no such modification it is labelled 'Unstressed'. However, sometimes the paper was ambiguous in what lines had added stress or the results from stressed and unstressed lines were pooled together, in this case we label it as 'Not Stated'. 
**g**: Hedge's g calculated using the compute.es package
**var.g**: The within study variance associated with the effect size, g
**mean/sd/n.low/high**: The means, standard deviation and sample size for the low or high sexual selection treatments, used to calculate lnCVR (meta-analysis of variance). Rows without these values had hedges g' derived from summary statistics (F, z, chi-square etc.)
**JIF**: Journal Impact factor at year of publication. Several impact factors were unable to be determined/found and are NA.

@@

# Make sure the factors are leveled in the same order as we write our prediction function (below)
restricted.data$Environment <- restricted.data$Environment %>% factor %>% relevel(ref="Unstressed")
restricted.data$Sex <- restricted.data$Sex %>% factor %>% relevel(ref="M")
restricted.data$Outcome.Class <- restricted.data$Outcome.Class %>% factor %>% relevel(ref="Indirect")

# run meta-analysis using the restricted data, and moderators
model.complete <- rma.mv(g, var.g, 
                         mods = ~ 1 + Sex + Environment + Taxon + Sex:Environment, 
                         random = ~ 1 | Study.ID/Outcome, 
                         method = "REML", 
                         data = restricted.data)

@@

# I cut all these out, since the trime-and-fill doesn't work, and those sub-models probably don't make sense
# trim.ambig <- trimfill.rma.uni(model.Ambiguous)
# funnel.rma(trim.ambig)
# regtest(model.Ambiguous, predictor = "vi")
# funnel.rma(model.Indirect)
# regtest(model.Indirect, predictor = "vi")
# funnel.rma(model.Direct)
# regtest(model.Direct, predictor = "vi")
# funnel.rma(model.complete2, type="rstandard", yaxis = "sei")
```

@@

  apatheme <- theme_bw() +  #My APA-format theme
    theme(panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          panel.border=element_blank(),
          axis.line=element_line(),
          text=element_text(family='Times'),
          legend.position='none')

@@

  #Now we can actually make the funnel plot.
  ggplot(dataset, aes(x = sqrt(var.g), y = g)) +
    geom_point(shape = 1) +
    xlab("Standard Error") + ylab("Effect size (Hedges' g)") +
    geom_line(aes(x = se.seq, y = ll95), linetype = 'dotted', data = dfCI) + # confidence lines
    geom_line(aes(x = se.seq, y = ul95), linetype = 'dotted', data = dfCI) +
    geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
    geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
    #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
    geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI, colour = "red") +
    geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI, colour = "red") +
    scale_x_reverse() +
    # scale_y_continuous(breaks = seq(-1.25,2,0.25)) + #Choose values that work for you based on your data
    coord_flip() +
    theme_classic()
    # apatheme
}

@@

```{r}

```{r, fig.width=10, fig.height=8}
prelim.data %>% ggplot(aes(x=JIF, y=(g)))+
  geom_jitter(color='darkgreen', alpha=.3, aes(size = (1/(var.g))))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'weight')
```

@@

####Time-lag Bias

@@

####Blinding

@@

```{r, fig.width=10, fig.height=8}
df.forest.model %>% ggplot(aes(x=Blinding, y=g))+
  geom_boxplot()+
  geom_jitter(aes(color=Blinding))+
  geom_hline(yintercept=0, linetype = 'dotted')
```

@@

We also collected sample sizes for each of the effect sizes calculated. Because we are dealing withj different taxa some studies are not suited to have sample sizes in the 1000's. We can simply inspect the sample size ande effect sizes through the following plot:
```{r, fig.width=5, fig.height=10}
prelim.data %>% ggplot(aes(x=(n), y = g))+
  geom_point()+
  xlim(0,2100)+
  facet_grid(Taxon~., scales='free')+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")
```

@@

####Generations

@@

```{r, fig.width=8, fig.height=7}
prelim.data %>% ggplot(aes(x=Generations, y=g, color=Taxon))+
  geom_point()+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")
```

>>>>

```{r warning=FALSE, message=FALSE, results='hide'}
library(knitr)
library(pander)
library(compute.es)
library(metafor)
library(dplyr)
library(lme4)
library(forestplot)
library(ggplot2)
library(MuMIn)
library(glmulti)
library(cowplot)
library(ggrepel)
library(reshape2)
library(RColorBrewer)
source("Vdodge_function.R") # nice function for ggplot
```

@@

**Study ID**: An ID given to the published paper the effect size is sourced from

**Group ID**: An ID given to the research group that may have published several papers on the same species usuing the same or very similar experimental setup

**Species and Species ID**: Same thing

**SS Strength, Ratios and SS Density's (Column 7-9)**: Various ratios of the number of males to females and the total number of individuals kept together in an experiment

**Ratio Category**: A three level category for the ratio of males to females (high, medium, low).

**Density Category**: A three level category for the density of males to females (high, medium, low).

**SSS.Categorical**: A three level category for the **density & the ratio** of males to females (high medium, low).

**Post cop and Pre cop**: Whether a study allowed Pre/Post-copulatory sexual selection (1) or not (0)
**Blind**: Whether the study was blind or not
**Generations** The number of generations run before fitness outcomes were measured
**Enforced Monogamy**: Whether the study had the low sexual selection treatment as enforced monogamy (YES) or not (NO). Not all studies compared enforced monogamy and SS+ treatments. Some used FB vs MB, where FB is the SS (low intensity). 
**Sex**: Whether the fitness outcome was measured for females (F), males (M) or both (B). Studies that reported 'both' would pool results from males and females.
**Ambiguous**: Is the fitness outcome ambiguous (YES) or not ambigous (NO). Ambiguous outcomes may be those that may not necessarily be directional, that is to say they may be a life history trait. 
**Outcome Class**: Grouped as Direct, Indirect or Ambiguous.
**Environment**: In the methods of the papers included in this study it was usually stated whether additional modifications to the experimental lines were made. Briefly, this was usually a modification that made conditions more stressful such as using a novel food source or elevated mutation load, the effect sizes from these experimental lines are labelled as 'Stressed'. If it was clearly stated that there was no such modification it is labelled 'Unstressed'. However, sometimes the paper was ambiguous in what lines had added stress or the results from stressed and unstressed lines were pooled together, in this case we label it as 'Not Stated'. 
**g**: Hedge's g calculated using the compute.es package
**var.g**: The within study variance associated with the effect size, g
**mean/sd/n.low/high**: The means, standard deviation and sample size for the low or high sexual selection treatments, used to calculate lnCVR (meta-analysis of variance). Rows without these values had hedges g' derived from summary statistics (F, z, chi-square etc.)
**JIF**: Journal Impact factor at year of publication. Several impact factors were unable to be determined/found and are NA.

@@

# Make sure the factors are leveled in the same order as we write our prediction function (below)
restricted.data$Environment <- restricted.data$Environment %>% factor() %>% relevel(ref="Unstressed")
restricted.data$Sex <- restricted.data$Sex %>% factor() %>% relevel(ref="M")
restricted.data$Outcome.Class <- restricted.data$Outcome.Class %>% factor() %>% relevel(ref="Indirect")
restricted.data$Taxon <- relevel(restricted.data$Taxon, ref = "Mouse")

model.complete <- rma.mv(g, var.g, 
                         mods = ~ 1 + Sex * Environment, # << ----- Add big model, then cull predictors to this one
                         random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome), 
                         method = "REML", 
                         data = restricted.data)

@@

# I cut all these out, since the trim-and-fill doesn't work, and those sub-models probably don't make sense
# trim.ambig <- trimfill.rma.uni(model.Ambiguous)
# funnel.rma(trim.ambig)
# regtest(model.Ambiguous, predictor = "vi")
# funnel.rma(model.Indirect)
# regtest(model.Indirect, predictor = "vi")
# funnel.rma(model.Direct)
# regtest(model.Direct, predictor = "vi")
# funnel.rma(model.complete2, type="rstandard", yaxis = "sei")
```

@@

  apatheme <- theme_bw() +  #My APA-format theme
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(),
          text = element_text(family = 'Times'),
          legend.position = 'none')

@@

  ggplot(dataset, aes(x = sqrt(var.g), y = g)) +
    geom_point(alpha = 0.4) +
    xlab("Standard Error") + ylab("Effect size (Hedges' g)") +
    geom_line(aes(x = se.seq, y = ll95), linetype = 'dotted', data = dfCI) + # confidence lines
    geom_line(aes(x = se.seq, y = ul95), linetype = 'dotted', data = dfCI) +
    geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
    geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
    #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
    geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI, colour = "tomato") +
    geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI, colour = "tomato") +
    scale_x_reverse() +
    # scale_y_continuous(breaks = seq(-1.25,2,0.25)) + #Choose values that work for you based on your data
    coord_flip() +
    theme_classic()
    # apatheme

@@

```{r, fig.width=10, fig.height=8}
prelim.data %>% ggplot(aes(x=log10(JIF), y=(g)))+
  geom_jitter(color='darkgreen', alpha=.3, aes(size = (1/(var.g))))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'weight')
```

@@

#### Time-lag Bias

@@

#### Blinding

@@

```{r, fig.width=10, fig.height=8}
df.forest.model %>% ggplot(aes(x=Blinding, y=g))+
  geom_boxplot()+
  geom_jitter(aes(color=Blinding))+
  geom_hline(yintercept=0, linetype = 'dotted') + 
  theme(legend.position = "none")
```

@@

We also collected sample sizes for each of the effect sizes calculated. Because we are dealing withj different taxa some studies are not suited to have sample sizes in the 1000's. We can simply inspect the sample size ande effect sizes through the following plot:
```{r, fig.width=5, fig.height=10}
prelim.data %>% ggplot(aes(x=n, y = g))+
  geom_point()+
  scale_x_log10(breaks = c(10, 100, 1000, 10000)) + 
  #xlim(0,2100)+
  facet_grid(Taxon~., scales='free')+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")
```

@@

#### Generations

@@

```{r, fig.width=8, fig.height=7}
prelim.data %>% ggplot(aes(x=Generations, y=g, fill=Taxon))+
  geom_point(shape=21, color = "grey20")+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed") + 
  scale_fill_brewer(palette = "Set3")
```
